Dysregulation of Cerebral Hemoglobin Content in Malawian Children with Malaria

### 2021-07-25 cm_follow_up_figures.Rmd
change order of variables in figures; export table to csv

# plot O2 sat
df_results_all_plot %>% 
  dplyr::filter(!grepl("avg|second", variable)) %>% 
  mutate(measurement = ifelse(grepl("conc", variable), "THC", "O2_sat")) %>% 
  mutate(measurement = factor(measurement, levels = c("THC", "O2_sat"))) %>% 
  mutate(pvals = ifelse(measurement == "THC", round(thc_fu_pval, 3), round(o2sat_fu_pval, 3))) %>%

  ggplot(aes(x = follow_up, y = value)) +
  geom_violin(aes(color = follow_up)) +
  geom_point(shape = 1) +
  stat_summary(fun = "median", geom = "crossbar", aes(x = follow_up, color = follow_up), size = 0.2, width = 0.5) +
  geom_path(aes(group = subject_id), lty = 2, alpha = 0.7) +

>>>>

# plot O2 sat
df_results_all_plot %>% 
  dplyr::filter(!grepl("avg|second", variable)) %>% 
  mutate(measurement = ifelse(grepl("conc", variable), "THC", "O2_sat")) %>% 
  mutate(measurement = factor(measurement, levels = c("O2_sat", "THC"))) %>% 
  mutate(pvals = ifelse(measurement == "THC", round(thc_fu_pval, 3), round(o2sat_fu_pval, 3))) %>%

  ggplot(aes(x = follow_up, y = value)) +
  geom_violin(aes(color = follow_up)) +
  geom_point(shape = 1, size = 2) +
  stat_summary(fun = "median", geom = "crossbar", aes(x = follow_up, color = follow_up), size = 0.2, width = 0.5) +
  geom_path(aes(group = subject_id), lty = 2, alpha = 0.7) +

### 2021-07-25 figures.Rmd
change order of variables in figures; export table to csv

  df_hbox <- read.csv("Data/master_datatable.csv") %>% 
    as_tibble() %>% 
    dplyr::select(-1) %>% 
    mutate(Status = factor(Status, levels = c("HV", "UM", "CM")))

@@

  dplyr::rename(" " = "variable",
                "Healthy volunteers" = "HV",
                "Uncomplicated malaria patients" = "UM",
                "Cerebral malaria patients" = "CM")

  # remove lactate data for HV
  df_clinical_res[11,]$`Healthy volunteers` <- "--"

@@

# select from the long df
df_hbox_voi_long <- df_hbox_long %>% dplyr::filter(variable %in% voi)

# create color vector
my_col <- c("CM" = "#00BFC4", "HV" = "#7CAE00", "UM" = "#F8766D")

@@

  ggplot(aes(x = Status, y = value)) +
  geom_violin(aes(color = Status)) +
  geom_jitter(position = position_jitter(0.2), shape = 1) +
  stat_summary(fun = "median", geom = "crossbar", aes(color = Status), size = 0.2, width = 0.5) +
  facet_wrap(~variable, scales = "free_y") +

@@

**Figure 1.** Shown are the variables measured by Oxiplex, plotted by patient status. Cerebral tissue hemoglobin concentration levels were different between cerebral malaria patients and the other two groups, but did not differ between uncomplicated malaria patients and healthy volunteers (Kruskal-Wallis adjusted p-value = 5.0x10^-7^; Dunn Test post-hoc adjusted p-values: CM vs HV = 3.2x10^-6^, CM vs UM = 2.8x10^-4^, HV vs UM = 0.87). Levels of cerebral hemoglobin O~2~ saturation did not differ among groups (Kruskal-Wallis adjusted p-value = 0.47).

@@

  ggplot(aes(x = Status, y = value)) +
  geom_violin(aes(color = Status)) +
  geom_jitter(position = position_jitter(0.2), shape = 1) +
  stat_summary(fun = "median", geom = "crossbar", aes(color = Status), size = 0.2, width = 0.5) +

@@

# include variables of interest, remove healthy volunteers from analysis

@@

# remove healthy volunteers for logistic regression analysis
df_hbox_log_reg <- df_hbox_impute %>% dplyr::filter(Status != "HV")

@@

collect_metrics(glm_res) %>% 
  dplyr::select(.metric, mean, std_err) %>% 
  mutate(mean = mean*100, std_err = std_err*100) %>% 
  mutate(Metric = c("Accuracy", "ROC AUC", "Sensitivity", "Specificity")) %>% 
  dplyr::select(c("Metric", "mean", "std_err")) %>%

  dplyr::rename("Mean" = "mean", "Standard error" = "std_err") %>% 
  mutate_if(is.numeric, round, 2) %>% 
  as.data.frame() %>%

  knitr::kable(align = rep('c', 4))

@@

hbox_final %>% 
  tidy(exponentiate = TRUE) %>% 
  mutate(estimate = ifelse(estimate < 10^-2, scientific(estimate, digits = 3), round(estimate, 2))) %>% 
  mutate(p.value = scientific(p.value, digits = 3),
         std.error = round(std.error, 2),
         statistic = round(statistic, 2)) %>% 
  dplyr::rename("Term" = "term", "Odds ratio" = "estimate", "Standard error" = "std.error", 
                "Statistic" = "statistic", "p-value" = "p.value") %>%

  as.data.frame() %>%
  knitr::kable(align = rep('c', 5))

@@

df_hbox %>% filter(Status != "HV") %>%

@@

# create run correlations of potential regressors
m_cor_hv <- df_hbox_impute %>% f_calc_cor(status = "HV")

@@

  # create plots
  m_cor_hv %>% f_plot_cor() + ggtitle("Healthy volunteers") + theme(legend.position = "none")

  m_cor_um %>% f_plot_cor() + ggtitle("Uncomplicated malaria patients") + theme(legend.position = "none")

  m_cor_cm %>% f_plot_cor() + ggtitle("Cerebral malaria patients") + theme(legend.position = "none")

>>>>

  df_hbox <- read.csv("Data/master_datatable.csv") %>% 
    as_tibble() %>% 
    dplyr::select(-1) %>% 
    mutate(Status = ifelse(Status == "HV", "HC", Status)) %>% 
    mutate(Status = factor(Status, levels = c("HC", "UM", "CM")))

@@

  dplyr::rename(" " = "variable",
                "Healthy children" = "HC",
                "Uncomplicated malaria" = "UM",
                "Cerebral malaria" = "CM")

  # remove lactate data for HC
  df_clinical_res[11,]$`Healthy children` <- "--"

  # export to CSV
  # write.csv(df_clinical_res, file = "Outputs/table_1.csv")

@@

# select from the long df
df_hbox_voi_long <- df_hbox_long %>% 
  dplyr::filter(variable %in% voi) %>% 
  mutate(variable = factor(variable, levels = c("avg_Hb_o2sat", "avg_Hb_conc",
                                                "Hb_o2sat_overall_alpha", "Hb_conc_overall_alpha",
                                                "Hb_o2sat_second_alpha", "Hb_conc_second_alpha")))

# create color vector
my_col <- c("CM" = "#00BFC4", "HC" = "#7CAE00", "UM" = "#F8766D")

@@

  ggplot(aes(x = Status, y = value)) +
  geom_violin(aes(color = Status)) +
  geom_jitter(position = position_jitter(0.2), shape = 1, size = 2) +
  stat_summary(fun = "median", geom = "crossbar", aes(color = Status), size = 0.2, width = 0.5) +
  facet_wrap(~variable, scales = "free_y") +

@@

**Figure 1.** Shown are the variables measured by Oxiplex, plotted by patient status. Cerebral tissue hemoglobin concentration levels were different between cerebral malaria patients and the other two groups, but did not differ between uncomplicated malaria patients and healthy volunteers (Kruskal-Wallis adjusted p-value = 5.0x10^-7^; Dunn Test post-hoc adjusted p-values: CM vs HC = 3.2x10^-6^, CM vs UM = 2.8x10^-4^, HC vs UM = 0.87). Levels of cerebral hemoglobin O~2~ saturation did not differ among groups (Kruskal-Wallis adjusted p-value = 0.47).

@@

  ggplot(aes(x = Status, y = value)) +
  geom_violin(aes(color = Status)) +
  geom_jitter(position = position_jitter(0.2), shape = 1, size = 2) +
  stat_summary(fun = "median", geom = "crossbar", aes(color = Status), size = 0.2, width = 0.5) +

@@

# include variables of interest, remove healthy children from analysis

@@

# remove healthy volunteers for logistic regression analysis
df_hbox_log_reg <- df_hbox_impute %>% dplyr::filter(Status != "HC")

@@

table2a <- collect_metrics(glm_res) %>% 
  dplyr::select(.metric, mean, std_err) %>% 
  mutate(mean = mean*100, std_err = std_err*100) %>% 
  mutate(Metric = c("Accuracy", "ROC AUC", "Sensitivity", "Specificity")) %>% 
  dplyr::select(c("Metric", "mean", "std_err")) %>%

  dplyr::rename("Mean" = "mean", "Standard error" = "std_err") %>% 
  mutate_if(is.numeric, round, 2) %>% 
  as.data.frame()

table2a %>% knitr::kable(align = rep('c', 4))

@@

table2b <- hbox_final %>% 
  tidy(exponentiate = TRUE) %>% 
  mutate(estimate = ifelse(estimate < 10^-2, scientific(estimate, digits = 3), round(estimate, 2))) %>% 
  mutate(p.value = scientific(p.value, digits = 3),
         std.error = round(std.error, 2),
         statistic = round(statistic, 2)) %>% 
  dplyr::rename("Term" = "term", "Odds ratio" = "estimate", "Standard error" = "std.error", 
                "Statistic" = "statistic", "p-value" = "p.value") %>% 
  as.data.frame()

table2b %>% knitr::kable(align = rep('c', 5))

```{r export_table2, eval = FALSE}

list(table2a, table2b) %>% write.csv("Outputs/table_2.csv")

@@

df_hbox %>% filter(Status != "HC") %>%

@@

# create run correlations of potential regressors
m_cor_hv <- df_hbox_impute %>% f_calc_cor(status = "HC")

@@

  # create plots
  m_cor_hv %>% f_plot_cor() + ggtitle("Healthy chilren") + theme(legend.position = "none")

  m_cor_um %>% f_plot_cor() + ggtitle("Uncomplicated malaria") + theme(legend.position = "none")

  m_cor_cm %>% f_plot_cor() + ggtitle("Cerebral malaria") + theme(legend.position = "none")
